---
title: "EDA Assignment Prep"
author: "Amber Camp"
date: "`r Sys.Date()`"
format: html
editor: visual
---

## EDA Assignment Prep: Pokemon data

### Libraries

```{r, message=FALSE}
library(here)
library(tidyverse)
library(janitor)
```

### Data

```{r, message=FALSE}
pokedata <- read_csv(here("data/Pokemon.csv"))

pokedata <- pokedata %>%
  clean_names()
```

### Describe the data: Pokemon

Data author: **ALBERTO BARRADAS**

<https://www.kaggle.com/datasets/abcsds/pokemon?resource=download>

Description: This is pokemon data for 721 pokemon and their stats.

How many observations (rows) are there? 800

How many variables (columns) are there? 13

### skimr: Quick overview

```{r, message=FALSE}
skim_data <- skimr::skim(pokedata)
```

### Summary: Quick overview

```{r, message=FALSE}
summary(pokedata)
```

With the summary, we can confirm what skimr told us: some of the data is in a class that will not be helpful for analysis. Let's take a closer look at changing classes below.

### Categorical vs. Continuous

```{r, message=FALSE, results='hide'}
class(pokedata$number) # checks class
pokedata$number <- as.factor(pokedata$number) # converts to factor
class(pokedata$number) # check class again
summary(pokedata$number) # review the data in this variable

# the reason I am converting to factor above is because I have looked at the data and decided
# that I want to treat the data as categorical rather than numeric. they are numbers, but
# they are id numbers related to each pokemon--not ordinal, measurements, or counts, etc. 

summary(pokedata$type_1)
pokedata$type_1 <- as.factor(pokedata$type_1)
class(pokedata$type_1)
summary(pokedata$type_1)

pokedata$name <- as.factor(pokedata$name)
pokedata$type_2 <- as.factor(pokedata$type_2)
pokedata$generation <- as.factor(pokedata$generation)

# here, i went ahead and converted the other variables to factor where it applies.
# you don't have to do this ahead. it can be done later.
# you can also convert back later at any point, if you need
```

How many categorical variables are there? 5 (maybe 6? legendary is "logical")

How many numeric variables are there? 7

What are some variables of interest to you? how many legendary?

------------------------------------------------------------------------

## EDA Assignment Prep: \[Insert Data Name Here\]

### Data

```{r, message=FALSE}

```

### Describe the data:

Data author:

Description:

How many observations (rows) are there?

How many variables (columns) are there?

### skimr: Quick overview

```{r, message=FALSE}

```

### Summary: Quick overview

```{r, message=FALSE}

```

### Categorical vs. Numeric

```{r, message=FALSE}

```

How many categorical variables are there?

How many numeric variables are there?

What are some variables of interest to you?

What kinds of questions could you explore with this data? i'm going to try again
